Equalizer 2.0 --\\Convergence of a Parallel Rendering Framework

### 2018-01-11 paperIEEE.tex
Tuning
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@76 db7c848b-7755-4df2-b763-8796797afa52

Virtual Reality is an important field for parallel rendering. It does however
require special attention to support it as a first-class citizen in a generic
parallel rendering framework. \textsf{Equalizer} has been used in many virtual
reality installations, such as the Cave2 (\cite{FNTTL:13}), the high-resolution
C6 CAVE at the KAUST visualization laboratory, and head-mounted displays
(\fig{FIG_teaser}). In the following we lay out the features needed to support
these installations. All features presented were motivated by application use
cases and have been validated with their respective users.

@@

To our knowledge, all parallel rendering systems have the focal plane coincide
with the physical display surface. For better viewing comfort, we introduce a
new dynamic focus mode, where the application defines the distance of the focal
plane from the observer, based on the current view direction of the
user. Initial experiments show that this is particularly effective for objects
placed within the immersive space, that is, in front of the physical displays.

@@

Traditional head tracking computes the left and right eye positions by using a
configurable interocular distance and the tracking matrix. However, human heads
are not symmetric, and by measuring individual users a more precise frustum can
be computed. \textsf{Equalizer} supports this through the optional configuration
(file or programmatically) of individual 3D translations for all eyes relative
to the tracking matrix.

@@

This feature allows applications to specify a scaling factor between the model
and the real world, to allow exploration of macroscopic or microscopic worlds in
virtual reality. The unit is per view, allowing different scale factors within
the same configuration. It scales both the specified projection surface as well
as the eye position (and therefore separation) to achieve the necessary effect.

@@

Applications can switch each view between mono and stereo rendering at runtime,
and can run both monoscopic and stereoscopic views concurrently
(\fig{FIG_teaser} (b)). While this is trivial for active stereo rendering, it
does potentially involve the start and stop of resources and processes for
passive stereo or stereo-dependent task decompositions (\sref{sec:reconfig}).

@@

The most obvious one is compression. Recently, many new compression algorithms
have been developed which exploit modern CPU architectures and deliver
compression rates well above one gigabyte per second. \textsf{Collage} uses the
Pression
library~\cite{pression}, which provides an unified interface for a number of
compression libraries, such as FastLZ~\cite{jesperfast}, Snappy~\cite{snappy}
and ZStandard~\cite{zstd}. It also contains a custom, virtually zero-cost RLE
compressor. Pression parallelizes the compression and decompression using
data decomposition. Note that this compression is generic, and implemented
transparently for the application. Applications can still use data-specific
compression and feed the compressed data to \textsf{Collage}, typically
disabling the
generic compression.

>>>>

Virtual Reality is an important field for parallel rendering. It does however
require special attention to support it as a first-class citizen in a generic
parallel rendering framework. \textsf{Equalizer} has been used in many virtual
reality installations, such as the Cave2 \cite{FNTTL:13}, the high-resolution C6
CAVE at the KAUST visualization laboratory, and head-mounted displays
(\fig{FIG_teaser}). In the following we lay out the features needed to support
these installations, motivated by application use cases.

@@

To our knowledge, all parallel rendering systems have the focal plane coincide
with the physical display surface. For better viewing comfort, we introduce a
new dynamic focus mode, where the application defines the distance of the focal
plane from the observer, based on the current view direction of the user.
Initial experiments show that this provides better viewing comfort, in
particular for objects placed in front of the physical displays.

@@

Traditional head tracking computes the left and right eye positions by using a
configurable interocular distance. However, human heads are not symmetric, and
by measuring individual users a more precise frustum can be computed.
\textsf{Equalizer} supports this through the optional configuration (file or
programmatically) of individual 3D translations for all eyes relative to the
tracking matrix.

@@

This feature allows applications to specify a scaling factor between the model
and the real world, to allow exploration of macroscopic or microscopic worlds in
virtual reality. The unit is per view, allowing different scale factors within
the same application. It scales both the specified projection surface as well
as the eye position (and therefore separation) to achieve the necessary effect.

@@

Applications can switch each view between mono and stereo rendering at runtime,
and run both monoscopic and stereoscopic views concurrently (\fig{FIG_teaser}
(b)). This switch does potentially involve the start and stop of resources and
processes for passive stereo or stereo-dependent task decompositions
(\sref{sec:reconfig}).

@@

The most obvious one is compression. Recently, many new compression algorithms
have been developed which exploit modern CPU architectures and deliver
compression rates well above one gigabyte per second. \textsf{Collage} uses the
Pression library~\cite{pression}, which provides an unified interface for a
number of compression libraries, such as FastLZ~\cite{jesperfast},
Snappy~\cite{snappy} and ZStandard~\cite{zstd}. It also contains a custom,
virtually zero-cost RLE compressor. Pression parallelizes the compression and
decompression using data decomposition. This compression is generic, and
implemented transparently for the application. Applications can also use
data-specific compression.
